Rapid evolution of a conspicuous dynamic visual signal in an introduced chameleon

### 2021-08-15 ms.Rmd
Fixed all text and refs. Need tio add one block of new analysis

@@

@@

@@

# Abstract
Rapid evolution frequently follows a significant change to an organism’s environment, such as during colonization of a novel habitat. We show that Jackson’s chameleons (*Trioceros j. xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) have evolved significantly more conspicuous sexual signals. Compared to Kenya, Oahu has few potential predators and therefore, very low predation pressure. We used visual modelling to quantify the conspicuousness of wild chameleons from both Hawaii and Kenya to multiple social and predator stimuli in order to test for character release of sexual signals. The strongest effect was for achromatic contrast, where Hawaiian chameleons were brighter during male contests and courtship and were more conspicuous to snake and bird predators. Chameleon signals also showed evidence of local adaptation. Relaxing natural selection can result in signal release in systems experiencing strong sexual selection, including animals with dynamic color change, and may act differently on the two primary visual channels (chromatic and achromatic).

# Introduction

When Darwin tackled the joint problems of speciation and evolutionary change through natural selection, he was of the view that evolution was a gradual process that took place over extended time periods @RN1. We now have numerous examples of evolutionary change following natural selection in both the laboratory and wild, and not only in invertebrates with short generation times. Vertebrates, including Caribbean *Anolis* lizards and Galapagos finches, have demonstrated a remarkable capacity for morphological change of functional traits (limb length, beak size) on ecological time scales (e.g. [@RN2; @RN3]). Furthermore, a far more recent development since Darwin’s seminal work on natural selection is the recognition that sexual selection can also drive significant changes in behavior and phenotype over short time intervals (e.g., [@RN4; @RN5; @RN6]).

Conspicuous color signals evolve because of the fitness advantage conferred to the signaller. These signals frequently become increasingly elaborate through a combination of male contest competition and female preference (e.g. runaway selection [@RN7; @RN8]). However, trait exaggeration is typically halted by natural selection because there is a threshold beyond which conspicuousness is too costly @RN9. Evolution can respond to this selection pressure in a variety of ways that include shifting conspicuous signals to areas of the body that might be concealed to predators. For example, many conspicuous lizards have colorful throats and venters that they flash at conspecific receivers but which are not easily visible to raptors, which are a significant predator of lizards @RN10. Conversely, many species will balance the costs of sexual and natural selection and this is the case in Trinidadian guppies where populations living in predator-free streams are far more conspicuous than in predator-rich streams [@RN6; @RN9]. These examples are from animals with static color signals. A parallel solution that reduces the risk of exposure to predators is dynamic color change because signals can be switched on and off relatively rapidly through neural or endocrine control [@RN11; @RN12]. A wide range of invertebrates (e.g. cephalopods) and vertebrates (e.g. lizards and frogs) have convergently evolved this ability [@RN13; @RN14; @RN15].

Unlike animals with static color signals, we have a poor understanding of how animals with dynamic color change respond to predation pressure and whether relaxing predation pressure will result in an increasingly conspicuous signal. Tackling this question in the wild is particularly difficult unless we can take advantage of unplanned changes to an organism’s environment such as during colonization of a novel habitat. Living organisms are occasionally translocated great distances by human action to environments far removed from their native range. These novel environments typically have a completely different suite of organisms, including predators, and offer a unique opportunity to study signal evolution in founder populations. Jackson’s chameleons (*Trioceros j. xanthopholus*) are just such a case (Fig.\@ref(fig:figure1) and fig. S1). In 1972, a shipment of chameleons was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade. The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island @RN16. On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors), while in Kenya they are preyed upon by a wide range bird, snake and occasionally, mammal, predators. This accidental introduction represents a unique opportunity to study the evolution of a dynamic visual signal by comparing native (Kenya) and introduced (Oahu) chameleons under low and high levels of natural selection.

Specifically, we tested the hypothesis that relaxed predation pressure results in character release of a dynamic visual signal. We tested the following predictions that underpin this hypothesis: 1.) display coloration used in male-male contests and courtship will be more conspicuous to a conspecific in Hawaii than Kenya; and 2.) display coloration used in response to an encounter with a bird and snake predator will be more conspicuous to both predator types in Hawaii than Kenya. We also asked whether color signals were consistent with local adaptation. Specifically, we predicted that chameleon display coloration in each population would be more conspicuous within its own signaling environment compared to the other population’s signaling environment. For example, Hawaiian chameleons were historically found in Kenya but under local adaptation they are predicted to be more conspicuous in their current signaling environment, Hawaii. Likewise, Kenyan chameleons have no history in Hawaii and are therefore predicted to be more conspicuous within their current signaling environment, Kenya (table S1).

To test for character release and whether color signals show local adaptation, we quantified the local signalling environment (fig. S2) and manipulated the social and predatory context chameleons encountered, in both their native Kenyan, and introduced Hawaiian, range (fig. S3). We presented wild-caught male chameleons with the following treatments: a male, a female, a model bird predator (African cuckoo-hawk, *Aviceda cuculoides*) and a model snake predator (boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1) and fig. S1,S4). We measured the spectral reflectance of males in dominant, subordinate, courtship, and antipredator (bird, snake) display states (Fig.\@ref(fig:figure2) and fig. S5). We then used visual modelling against a known chameleon visual system for sexual signals, and a bird and snake visual system for predation trials, to quantify chromatic and achromatic contrast against the backgrounds of each environment (supplementary text).

Under relaxed predation, sexual selection is expected to favor more conspicuous dynamic sexual signals in Hawaii than Kenya.  In support of this prediction, male display coloration exhibited in social contexts (i.e., male-male contests and courtship) was significantly more conspicuous with respect to brightness (Fig.\@ref(fig:figure2) &  Fig.\@ref(fig:figure3)A) to a conspecific in Hawaii compared with Kenya (male-male contests: achromatic contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: achromatic contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`), but not significantly different in chromatic contrast (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`), and this was true across all body regions measured (See supplementary analyses). Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (achromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).

As a consequence of signal release, sexual signals are expected to be more easily detected by predators (i.e., snakes and birds) in Hawaii. Also, differential predation pressure resulting from differences in predation rates by snake and bird predators would also result in populations varying in their conspicuousness to each predator type.  In support of these predictions, Hawaiian chameleons were significantly more conspicuous to both snake and bird predators with respect to brightness (Fig.\@ref(fig:figure3)C; achromatic contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`). Differences in chromatic contrast between populations, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`). Interestingly, Kenyan and Hawaiian chameleons were equally detectable by snakes, but Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; Birds (Hawaii-Kenya): `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`).

Local adaptation of sexual signals would result in Hawaiian chameleons being more conspicous to conspecifics in the Hawaiian environment compared to their native Kenyan environment. While it was not possible to conduct common garden experiments to provide definitive tests of local adaptation, our results support the local adaptation hypothesis. Hawaiian chameleons were more conspicuous to other chameleons against heir own (Hawaiian) background as opposed to the Kenyan background, in both social contexts and for both achromatic and chromatic contrast (Fig.\@ref(fig:figure4); Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: achromatic ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: achromatic ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`).

We have documented a surprisingly strong effect of environment on the evolution of a conspicuous, dynamic sexual signal. For a founding population of chameleons, the tropical environment of Hawaii represents a particularly favorable set of conditions with respect to habitat, temperature, and rainfall. Furthermore, Jackson’s chameleons are dietary generalists @RN17 and likely experience low competition for relatively abundant arthropod prey. But most importantly, Hawaii has no snakes and no raptors that feed on lizards. Domestic and feral cats would be their main threat on Oahu, and cats have very limited color vision @RN18. In the absence of the usual suite of predators they would experience in their native range of Kenya—a diverse community of snake and bird predators, the conditions were highly favorable for character release of sexual signals—particularly in a polygynous species in which males have intense male contest competition and elaborate courtship displays (Fig.\@ref(fig:figure1)). Not only do males use colorful displays, but they have three horns that they lock and twist into rivals during intense bouts that sometimes result in physical injury. Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated—an increasingly bright sexual signal used in both male contests and during courtship.

Evolutionary biologists, particularly those interested in speciation, seldom consider the role of natural selection in shaping sexually selected signals even though sexual signals evolve in an ecological context @RN19 (but see [@RN5; @RN20, @RN21]). Landmark studies such as those on Trinidadian guppies [@RN6; @RN9]  have shown the important interaction between natural and sexual selection in static signals following experimental manipulation. Ours is one of the only studies that show that a dynamic sexual signal—rapid color change—is likewise constrained by natural selection. Our results suggest that dynamic sexual signals are not simply plastic in a way that allows an individual to moderate the extent to which they ‘turn on’ a display but rather, that there is a threshold constrained by natural selection. This may in part be explained by the nature of color change in chameleons. Chameleons have two layers of superposed iridiophores that facilitates color change that produces both camouflage and conspicuous sexual signals @RN22. Interestingly, while Hawaiian chameleons have evolved brighter display coloration, other elements of color including hue and saturation have not experienced similar levels of selection except to become more conspicuous against their own background (local adaptation). This suggests that the two visual channels (achromatic vs chromatic) experience differing levels of selection which is perhaps not surprising given that they act independently of one another @RN23. While studying evolution is challenging because it is not possible to go back in time to witness key evolutionary events, we can take advantage of unique opportunities such as an invasion, to better understand how natural and sexual selection interact to shape animal signals.

```{r figure1, cache=FALSE, fig.cap = "Chameleon color signal change in response to different social and predatory stimuli. Male chameleons experience intense sexual selection. During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal. They also readily fight by locking horns and sometimes pierce their rivals skin with their horns. (A) A typical dominant male in display coloration. (B) A subordinate male that lost a contest and turned from bright yellow to brown. (C) Two males fighting, both are in display colouration and at a stage where they are relatively even matched. (D) A male courting female in full display color while the female has turned to a contrasting colour, which suggests rejection of the male"}
library(magick)

@@

```{r figure2, fig.width = 14, fig.height= 14, fig.cap="Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya. The context for measurement was (A) male contest displays; (B) courtship; (C) bird predator; and (D) snake predator. More details in online supplementary material and Fig. S5"}
library(magick)
img <- image_read_pdf("./output/figures/Figure3.pdf")
img

@@

```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "Chromatic and achromatic contrast in Just Noticiable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves). (A & B) Acromatic and chromatic contrast of male chameleons during male-male contests and female courtship. JNDs are calculated based on the chameleon visual system. (B & C) Acromatic and chromatic contrasts of male chameleons during snake and bird predator encounters. JNDs are calculated based on the snake and bird visual systems. Contrasts between means, 'Beta', that are bold indicate significant effects."}
rerun = FALSE
if(rerun){
## Make figures that average across body regions, but display Social Context

@@

```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "Evidence for local adaptation of color signals in both (A) chromatic- and (B) achromatic contrast for Hawaiian chameleons. For local adaptation, signals are predicted to be more conspicuous against their own background. In this case, Hawaiian chameleons were both significantly more chromatic and brighter (achromatic contrast)  against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
rerun=FALSE

>>>>

# Sample sizes
N_display <- data.frame(Display_1 %>% group_by(Population) %>% summarise(n = length(unique(ID))))

@@

# Sample sizes
N_Courtship <- data.frame(Courtship %>% group_by(Population) %>% summarise(n = length(unique(ID))))

@@

# Sample sizes
N_Pred <- data.frame(Predator %>% group_by(Population, Predator) %>% summarise(n = length(unique(ID)))) %>% arrange(Predator)

@@

# Summary
Colonization of novel environments can free introduced populations from the constraints imposed by native-range predators and promote rapid evolutionary change. We demonstrate such character release for color signals in Jackson’s chameleons (*Trioceros j. xanthopholus*) inadvertently introduced from Kenya to Hawaii (Oahu) where there are few lizard predators. Hawaiian chameleons displayed more conspicuous color signals than Kenyan chameleons during male contests and courtship, were less cryptic in response to bird and snake predators, and showed greater change between display and anti-predator color states. Consistent with local adaptation, the display colors of Hawaiian chameleons were more conspicuous in their local than ancestral habitats. These results demonstrate that relaxed predation pressure can result in character release of dynamic social signals in introduced species experiencing strong sexual selection.

# Main text

When species encounter novel environments, such as during human-mediated introductions, evolutionary change can occur more rapidly than Darwin ever thought possible [@RN1; @RN2; @RN3; @RN4]. In particular, release from predation by native-range predators enables evolutionary exploration of entirely new regions of the adaptive landscape [@RN4; @RN5]. Such character release should be most evident for social and sexual signals; released from constraints on conspicuousness imposed by predation risk, signals could elaborate in many directions [e.g., @RN4; @RN6; @RN7; @RN8]. However, we are only beginning to understand how social and sexual signals change in response to rapid environmental change and there remain very few empirical examples from natural populations [but see @RN7; @RN9; @RN10]. This is especially true for species with dynamic color change, for which we are not aware of any examples. Here, we examine character release of a dynamic color signal by taking advantage of an unintended evolutionary ‘experiment’: accidental translocation of chameleons to a novel, low predation environment.

In 1972, a shipment of Jackson’s chameleons (*Trioceros jacksonii xanthopholus*; (Fig.\@ref(fig:figure1) and fig. S1)) was sent from Kenya to the Hawaiian island of Oahu, destined for the pet trade. The animals arrived in poor condition and were left outdoors to recover, at which point they dispersed and became established on the island [@RN11]. Jackson’s chameleons have a high intrinsic population growth rate (9--12 month generation time and live birth of up to 50 young @RN24), enabling their rapid establishment and potentially rapid evolution (approx. 50 -- 65 generations). On Oahu, the number of potential predators of chameleons is very few (no snakes or lizard-eating raptors, other potential bird predators absent or rare, supplementary text), while in Kenya they are preyed upon by a wide range of bird, snake and occasionally, mammal, predators. This accidental introduction enables us to test whether relaxed predation results in character release of a dynamic visual signal and whether this might be explained by local adaptation.

Chameleons are famous for their rapid color change [@RN12; @RN13; @RN14]. When presented with another chameleon, male Jackson’s chameleons adopt a characteristic display posture and become intensely yellow-green to signal dominance or during courtship (Fig. 1, fig. S1, S3). At other times, they are a duller green and can darken to brown or become mottled green and brown, particularly in response to perceived predation risk (Fig.\@ref(fig:figure1), fig. S4A). Their different color states can appear more or less conspicuous depending on the context, background, and viewer (e.g., conspecific or predator). If introduced chameleons in Hawaii experience character release from reduced predation, we expect Hawaiian chameleons more conspicuous male display colors in response to conspecifics as well as less cryptic coloration in response to bird and snake predators. We should also see evidence of local adaptation: color signals should be more conspicuous to conspecifics against the local than ancestral background.

To test these predictions, we presented wild-caught male chameleons in both Hawaii and Kenya with either a male chameleon, a female chameleon, a bird predator (stuffed African cuckoo-hawk, *Aviceda cuculoides*) or a snake predator (replica boomslang, *Dispholidus typus*) (Fig.\@ref(fig:figure1), fig. S1, S4). We measured the spectral reflectance of male color states during male-male display (dominant in male-male contests, N = `r N_display[1,2]` Hawaii, N = `r N_display[2,2]` Kenya), courtship (N = `r N_Courtship[1,2]` Hawaii, N = `r N_Courtship[2,2]` Kenya) and in response to bird (N = `r N_Pred[1,3]` Hawaii, N = `r N_Pred[2,3]` Kenya) and snake (N = `r N_Pred[3,3]` Hawaii, N = `r N_Pred[4,3]` Kenya) predators (Fig.\@ref(fig:figure2), fig. S5). We then modelled how conspicuous these colors would appear to the relevant receiver’s visual system (chameleon, bird, or snake), estimated as chromatic or luminance contrast (Just Noticeable Differences, JNDs^@RN25 against the background vegetation of each environment (Hawaii or Kenya; supplementary text).

In support of predictions of character release, male display coloration of Hawaiian chameleons had higher luminance contrast against the local background than those of Kenyan chameleons (Fig.\@ref(fig:figure2), Fig.\@ref(fig:figure3)A) (male-male contests: luminance contrast ($dl$); $F_{`r Population_ConspicDisplayAC$df_num`,`r Population_ConspicDisplayAC$df_denom`}$ = `r Population_ConspicDisplayAC$Fstat`, p = `r p_value(Population_ConspicDisplayAC$p_val)`; courtship displays: luminance contrast ($dl$); $F_{`r Population_ConspicCourtshipAC$df_num`,`r Population_ConspicCourtshipAC$df_denom`}$ = `r Population_ConspicCourtshipAC$Fstat`, p = `r p_value(Population_ConspicCourtshipAC$p_val)`). This result was consistent for all body regions measured (no significant interaction with body region; supplementary text). However, Hawaiian chameleons did not differ from Kenyan chameleons in the chromatic contrast of color signals (Fig.\@ref(fig:figure3)B, male-male contests: $F_{`r Population_ConspicDisplayCC$df_num`,`r Population_ConspicDisplayCC$df_denom`}$ = `r Population_ConspicDisplayCC$Fstat`, p = `r p_value(Population_ConspicDisplayCC$p_val)`; courtship: $F_{`r Population_ConspicCourtshipCC$df_num`,`r Population_ConspicCourtshipCC$df_denom`}$ = `r Population_ConspicCourtshipCC$Fstat`, p = `r p_value(Population_ConspicCourtshipCC$p_val)`).

The importance of luminance contrast is consistent with the nature and mechanism of rapid color change in individual chameleons, which usually involves much greater luminance than chromatic change (Fig.\@ref(fig:figure1)). Specifically, color change is caused by dispersion or concentration of melanin within melanophore pigment cells @RN15 or changes in the spacing of guanine crystals within iridophores @RN14, both of which strongly affect luminance. Accordingly, character release was observed in the overall luminance of signals rather than their hue.

Released from predation, Hawaiian chameleons were less cryptic than Kenyan chameleons when threatened by both bird and snake predators. Hawaiian chameleons had higher luminance contrast against the local background to the visual system of the corresponding predator (Fig.\@ref(fig:figure3)C; luminance contrast: ($dl$); $F_{`r Population_AC_Pred$df_num`,`r Population_AC_Pred$df_denom`}$ = `r Population_AC_Pred$Fstat`, p = `r p_value(Population_AC_Pred$p_val)`, no interaction between population and predator type). Differences between populations in chromatic contrast against the background, however, depended on the predator (significant predator by population interaction: $F_{`r back_pred_CC_Pred$df_num`,`r back_pred_CC_Pred$df_denom`}$ = `r back_pred_CC_Pred$Fstat`, p = `r p_value(back_pred_CC_Pred$p_val)`). Hawaiian chameleons were significantly more conspicuous to birds compared to Kenyan chameleons (Fig.\@ref(fig:figure3)D; Birds (Hawaii-Kenya): $\Beta$ = `r CI_hyp_test[2,2]`, 95% CI = `r CI_hyp_test[2,3]` to `r CI_hyp_test[2,4]`, p = `r CI_hyp_test[2,8]`, JND values > 5). By comparison, chameleons from both populations had very low and similar chromatic contrast to the visual system of snakes (Fig.\@ref(fig:figure3)D; Snakes (Hawaii-Kenya): $\Beta$ = `r CI_hyp_test[1,2]`, 95% CI = `r CI_hyp_test[1,3]` to `r CI_hyp_test[1,4]`, p = 1; JND values ~ 2). The much lower chromatic contrast of chameleons to the visual system of snakes than birds reflects the much poorer chromatic discrimination of trichromatic snakes compared to tetrachromatic birds [@RN18]. The poor color discrimination of snakes may also explain why chameleons from Hawaii show little change relative to their Kenyan counterparts in their chromatic response to snakes, despite the absence of snakes on Hawaii.

Character release may reflect rapid evolutionary change, phenotypic plasticity or a combination of these processes [@RN19; @RN20]. Rapid evolutionary change results in local adaptation, whereby signals should become more conspicuous to conspecifics in their local environment than in ancestral environments. Consistent with local adaptation, we found that Hawaiian chameleons were more conspicuous to other chameleons against their own (Hawaiian) background than against their ancestral Kenyan background, in both social contexts, and for both luminance and chromatic contrast (Fig.\@ref(fig:figure4); Contrast between Hawaiian lizard on Kenyan vs Hawaiian on Hawaiin background - male-male contests: luminance ($dl$), $\beta$ = `r CI_dis_AC[1,1]`, 95% CI = `r CI_dis_AC[1,2]` to `r CI_dis_AC[1,3]`, p = `r p_displ_AC`, chromatic ($dS$) $\beta$ = `r CI_Disp_CC[1,1]`, 95% CI = `r CI_Disp_CC[1,2]` to `r CI_Disp_CC[1,3]`, p = `r p_displ_CC`; courtship: luminance ($dl$), $\beta$ = `r CI_AC_Courtship[1,1]`, 95% CI = `r CI_AC_Courtship[1,2]` to `r CI_AC_Courtship[1,3]`, p = `r p_courtship_AC`, chromatic ($dS$) $\beta$ = `r CI_CC_Courtship[1,1]`, 95% CI = `r CI_CC_Courtship[1,2]` to `r CI_CC_Courtship[1,3]`, p = `r p_courtship_CC`). Although these results are consistent with local adaptation, this does not exclude a role for phenotypic plasticity, which can also evolve rapidly [@RN26].

To gauge the extent of color plasticity, we examined individual change between display (male-male contest) and anti-predator (bird) color states .........

We have documented a surprisingly strong effect of recent release from predation on a conspicuous, dynamic visual signal. Oahu has no snakes or raptors that feed on lizards. Domestic and feral cats would be their main threat on Oahu, and cats have very limited color vision [@RN22]. In the absence of the usual suite of predators they would experience in their native range of Kenya -- a diverse community of snake and bird predators -- the conditions were highly favorable for character release of social signals. Jackson’s chameleons likely experience strong sexual selection, given their polygynous mating system, elaborate courtship displays and intense male contest competition (Fig.\@ref(fig:figure1)) in which males lock and twist their three horns into rivals during intense bouts that sometimes result in physical injury. Consequently, a ‘perfect storm’ of conditions set the stage for what we have demonstrated -- character release of a dynamic social signal over a short ecological time scale.

Rapid color change enables animals to be highly conspicuous during social interactions and highly cryptic at other times [@RN12]. Our results suggest that even for species capable of dynamic color change there is an upper threshold to signal intensity that is constrained by natural selection. This novel finding raises the question of whether other animals that have convergently evolved dynamic color signals, such as cephalopods, frogs, other lizards, and fishes, may be likewise constrained. More generally, our study highlights the opportunities provided by translocations and invasions, which are often otherwise detrimental, to study natural and sexual selection in the wild.

## OLD

 Kenyan chameleons, from a population where predation pressure is much higher, were significantly more crypic compared to their introduced Hawaiian counterparts (achromatic contrast between Hawaiian and Kenyan Populations, male-male contests: $\beta$ = `r robust_ConspicDisplayAC$beta[3]`, 95% CI = `r robust_ConspicDisplayAC_CI$CI_L[3]` to  `r robust_ConspicDisplayAC_CI$CI_U[3]`, df = `r robust_ConspicDisplayAC$df[3]`, t =  `r robust_ConspicDisplayAC$tstat[3]`, p = `r p_value(robust_ConspicDisplayAC$p_Satt[3])`, courtship: $\beta$ = `r robust_ConspicCourtshipAC_CI$beta[3]`, 95% CI = `r robust_ConspicCourtshipAC_CI$CI_L[3]` to  `r robust_ConspicCourtshipAC_CI$CI_U[3]`, df = `r robust_ConspicCourtshipAC_CI$df[3]`, t =  `r summary(ConspicCourtshipAC)$coefficients[3, "t value"]`, p = `r p_value(robust_ConspicCourtshipAC$p_Satt[3])`, see Fig.\@ref(fig:figure3)).

```{r figure1, cache=FALSE, fig.cap = "**Chameleon color signal change in response to different social and predatory stimuli. Male chameleons experience intense sexual selection**. During the breeding season they change from a neutral state (green) to a highly conspicuous bright yellow display signal. They also readily fight by locking horns and sometimes pierce their rival’s skin with their horns. (A) A typical dominant male in display coloration. (B) A subordinate male that lost a contest and turned from bright yellow to brown. (C) Two males fighting, both are in display coloration and at a stage where they are relatively even matched. (D) A male courting female in full display color while the female has turned to a contrasting color, which suggests rejection of the male"}
library(magick)

@@

```{r figure2, fig.width = 14, fig.height= 14, fig.cap="**Mean spectral reflectance curves for male chameleons for representative body regions (gular, top flank) and background (leaves) for Hawaii and Kenya**. The context for measurement was (A) male contest displays; (B) courtship; (C) bird predator; and (D) snake predator. More details in online supplementary text and Fig. S5"}
library(magick)
img <- image_read_pdf("./output/figures/Figure3.pdf")
img

@@

```{r figure3, cache = FALSE, fig.width = 14, fig.height= 14, fig.cap= "**Chromatic and luminance contrast in Just Noticeable Differences (JND), of Hawaiian and Kenyan chameleons against their respective backgrounds (i.e., average environment of stems and leaves)**. (A & B) Luminance and chromatic contrast of male chameleons during male-male contests and female courtship. JNDs are calculated based on the chameleon visual system. (B & C) Luminance and chromatic contrasts of male chameleons during snake and bird predator encounters. JNDs are calculated based on the snake and bird visual systems. Contrasts between means, ‘Beta’, that are bold indicate significant effects.."}
rerun = FALSE
if(rerun){
## Make figures that average across body regions, but display Social Context

@@

```{r figure4, cache = FALSE, fig.width = 14, fig.height= 9, fig.cap= "**Evidence for local adaptation of color signals in both (A) chromatic- and (B) luminance contrast for Hawaiian chameleons**. For local adaptation, signals are predicted to be more conspicuous against their own background. In this case, Hawaiian chameleons were both significantly more chromatic and brighter (luminance contrast) against their own background compared to the Kenyan background suggesting that selection has acted on increasing local signal conspicuousness."}
rerun=FALSE
